Master's Project Journal

### 2019-04-17 thesis/FPGABasedADPLLs.tex
? pages

\subsection{SigNum Detector}
The SigNum detector is based on the methodology used in the Bang-Bang \acl{PD} seen previously, but rather than just performing sign detection, the magnitude of the phase difference can also be measured. In the case of the \ac{FPGA} clocked design the counter handled the representation of both sign and magnitude, but as the name SigNum (Sign Number) suggested both are handled separately in this design. The modified Bang-Bang detector establishes which signal occurred first and generates a control signal that is analogous to the time delay between a rising edge on each input signal. A \ac{TDL} then uses this control signal to convert the phase difference between the signals to a digital signal. This magnitude is combined with the sign detected by the bang-Bang detector to form a two's complement error signal.

>>>>

\subsection{SigNum Detector}\label{section:signum}
The SigNum detector is based on the methodology used in the Bang-Bang \acl{PD} seen previously, but rather than just performing sign detection, the magnitude of the phase difference can also be measured. In the case of the \ac{FPGA} clocked design the counter handled the representation of both sign and magnitude, but as the name SigNum (Sign Number) suggested both are handled separately in this design. The modified Bang-Bang detector establishes which signal occurred first and generates a control signal that is analogous to the time delay between a rising edge on each input signal. A \ac{TDL} then uses this control signal to convert the phase difference between the signals to a digital signal. This magnitude is combined with the sign detected by the bang-Bang detector to form a two's complement error signal.

### 2019-04-17 thesis/NetworkImplemenation.tex
? pages

\subsubsection{Phase Detector}
As previously mentioned the \ac{PFD} is also implemented using the \ac{FPGA} clocked approach, with sign detection carried out using a Moore Machine. The state transition diagram given as the example in Chapter \ref{chap:3} is that used to design this sign detector. 
\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{../state_trans_new}
	\caption[Example State Transition Diagram for a Moore Machine]{Example State Transition Diagram for a Moore Machine.}
	\label{fig:state_trans_reprint}
\end{figure}

@@

\subsection{\acs{ADPLL} Design 2}
The second \ac{ADPLL} design implemented during the course of this project was somewhat of a stepping stone between fully synchronous and asynchronous to the \ac{FPGA} clock. The clocked phase detector is reused from above, albeit with an altered error width, while the \ac{DCO} is replaced by a \acl{RO}. While the phase detector was known to work due to testing in \ac{ADPLL} 1, the \ac{DCO} could not be simulated using accurate timings so all testing had to be done in hardware. This presented a challenge, as depending on the implementation, the characteristics of the \ac{DCO} could change, most importantly the centre frequency.

@@

@@

\begin{table}[!h]%TODO numbers!
	\begin{center}
		\begin{tabular}{|l|r|r|r|}
			\multicolumn{4}{c}{\ac{DCO} Tuning Range} \T\\
			\hline
			\multicolumn{1}{|c|}{-} & \multicolumn{1}{c|}{Minimum} & \multicolumn{1}{c|}{Step} & \multicolumn{1}{c|}{Maximum} \T\\
			\hline
			Frequency & $4.571~\si{\mega\hertz}$ & $29.180~\si{\kilo\hertz}$ (at $5~\si{\mega\hertz}$) & $5.518~\si{\mega\hertz}$ \T\\
			\hline
			Period & $218.2~\si{\nano\second}$ & $1.176~\si{\nano\second}$ & $181.1~\si{\nano\second}$ \T\\
			\hline
		\end{tabular}
		\begin{tabular}{|l|r|l|r|}
			\multicolumn{4}{c}{Configuration as Implemented} \T\\
			\hline
			\ac{RO} num. of inverters (max) & 373 & \ac{RO} Control Width & 5 bits \T\\
			\hline
			\ac{RO} Bias Point & 16 & Runtime Gains & Enabled \T\\
			\hline
			Phase Detector Error Width & 5 & Error Sum Weight Width & 4 \T\\
			\hline
			\acs{kp} & $2^{-7}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-9}\times[0,15]\cap\mathbb{Z}$ \T\\
			\hline
			Detection Resolution & $3.875~\si{\nano\second}$ & Detection Phase Resolution & $6.975\si{\degree}$ (at $5~\si{\mega\hertz}$)\\
			\hline
		\end{tabular}
	\end{center}
	\caption[ADPLL Design 2 Summary]{ADPLL Design 2 Summary.}
	\label{table:adpll2}
\end{table}

\subsection{\acs{ADPLL} Design 3}
\subsubsection{\acl{DCO}}
\subsubsection{Phase Detector}

\begin{figure}[h]
	\centering
	\subfloat[Modified Bang-Bang sign detector.]{{\includegraphics[width=0.8\textwidth]{../new_pdet1} }}\\
	\subfloat[\acl{TDL}.]{{\includegraphics[width=1.0\textwidth]{../new_pdet2} }}%
	\caption[\acl{PD} RTL diagrams]{\acl{PD} RTL diagrams.}
	\label{fig:signum_impl}
\end{figure}

\section{Measurement Setup}\label{section:measurement_setup}

\section{\acs{ADPLL} Characterisation}

>>>>

\subsubsection{Phase Detector}
As previously mentioned the \ac{PFD} is also implemented using the \ac{FPGA} clocked approach, with sign detection carried out using a Moore Machine. The state transition diagram given as the example in Chapter \ref{chap:3} is that used to design this sign detector. The hardware description of this block is found in \textsc{PhaseDetector.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:PhaseDetector}.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{../state_trans_new}
	\caption[Example State Transition Diagram for a Moore Machine]{Example State Transition Diagram for a Moore Machine.}
	\label{fig:state_trans_reprint}
\end{figure}

@@

\subsection{\acs{ADPLL} Design 2}
The second \ac{ADPLL} design implemented during the course of this project was somewhat of a stepping stone between fully synchronous and asynchronous to the \ac{FPGA} clock. The clocked phase detector is reused from above, albeit with an altered error width, while the \ac{DCO} is replaced by a \acl{RO}. While the phase detector was known to work due to testing in \ac{ADPLL} 1, the \ac{DCO} could not be simulated using accurate timings so all testing had to be done in hardware. This presented a challenge, as depending on the implementation, the characteristics of the \ac{DCO} could change, most importantly the centre frequency. The module implementing this \ac{ADPLL} can be found in \textsc{NetworkRingADPLL.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:network_ring_adpll}.

@@

The module implementing the \ac{RO} can be found in \textsc{RingOsc.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:ring_osc}.

@@

\begin{table}[!h]%TODO redo numbers!
	\begin{center}
		\begin{tabular}{|l|r|r|r|}
			\multicolumn{4}{c}{\ac{DCO} Tuning Range} \T\\
			\hline
			\multicolumn{1}{|c|}{-} & \multicolumn{1}{c|}{Minimum} & \multicolumn{1}{c|}{Step} & \multicolumn{1}{c|}{Maximum} \T\\
			\hline
			Frequency & $4.571~\si{\mega\hertz}$ & $29.180~\si{\kilo\hertz}$ (at $5~\si{\mega\hertz}$) & $5.518~\si{\mega\hertz}$ \T\\
			\hline
			Period & $218.2~\si{\nano\second}$ & $1.176~\si{\nano\second}$ & $181.1~\si{\nano\second}$ \T\\
			\hline
		\end{tabular}
		\begin{tabular}{|l|r|l|r|}
			\multicolumn{4}{c}{Configuration as Implemented} \T\\
			\hline
			\ac{RO} num. of inverters (max) & 373 & \ac{RO} Control Width & 5 bits \T\\
			\hline
			\ac{RO} Bias Point & 16 & Runtime Gains & Enabled \T\\
			\hline
			Phase Detector Error Width & 5 & Error Sum Weight Width & 4 \T\\
			\hline
			\acs{kp} & $2^{-7}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-9}\times[0,15]\cap\mathbb{Z}$ \T\\
			\hline
			Detection Resolution & $3.875~\si{\nano\second}$ & Detection Phase Resolution & $6.975\si{\degree}$ (at $5~\si{\mega\hertz}$)\\
			\hline
		\end{tabular}
	\end{center}
	\caption[\ac{ADPLL} Design 2 Summary]{\ac{ADPLL} Design 2 Summary.}
	\label{table:adpll2}
\end{table}

\subsection{\acs{ADPLL} Design 3}
\acs{ADPLL} Design 3 has no dependence on the clock provided by the \ac{FPGA}'s clock management utility, as both the \ac{RO} and the \ac{TDC} are designed using inverters, while a Bang-Bang detector replaces the \ac{FPGA} clock reliant \ac{FSM} of the two previous designs. This introduces a second source of implementation based variation between \ac{ADPLL}s, however, this has significantly lesser impact on the frequency range of the oscillator and no impact on the centre frequencies. The lack of an \ac{FPGA} clocked element results means this design will have the closest behaviour to that of an \ac{ASIC} based \ac{ADPLL} implementation, although as the level of variation due to implementation is much larger on an \ac{FPGA}, the impact of non-idealities will be exaggerated.
%same interface
In order to allow quick exchange of \ac{PD}s or other modules, the module in which this \ac{ADPLL} is implemented is still retains the \ac{FPGA} clock as an input. In fact, this module shares its implementation with that of \ac{ADPLL} 2, as while the \ac{PD} designs are drastically different, the interface is identical. The module implementing this \ac{ADPLL} can be found in \textsc{NetworkRingADPLL.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:network_ring_adpll}.

\subsubsection{\acl{DCO}}
The \ac{DCO} used in \ac{ADPLL} 3 is identical in form to that used in \ac{ADPLL} 2, so the tuning range, period step and all effects of implementation variations carry over to this \ac{ADPLL}. The decision to leave the \ac{RO} unmodified is an easy one, as the intended centre frequency is the same and changing the control code width would remove the ability to fix the cells of each \ac{RO}. Fixing the cells is beneficial as it will maintain the oscillator layout between implementation ``runs'', and thus allow for a better comparison between designs 1 \& 2.

\subsubsection{Phase Detector}
Instead of an \ac{FPGA} clocked phase detector, \ac{ADPLL} 3 uses the modified Bang-Bang detector and \ac{TDL} approach to phase comparison seen in Section \ref{section:signum}. Sign detection is implemented exactly as described in that section, with the double \ac{SR} latch approach used to determine the sign of the phase error. A flip-flop is inserted after the second \ac{SR} latch, clocked on the signal \texttt{done}, which holds the output value and ensures that the sign changes at the same time as the magnitude. The previous discussion of this phase detector proposed no implementation of the edge detectors, which are easily implemented using flip-flops clocked on the signal of interest, with a constant $1$ on the $D$ input. In this form the edge detector is single use only, as subsequent edges will not cause any change in the flip-flop output. This is important as it ensures the measurement interval will only terminate when a rising edge is seen on the signal that did not start the measurement interval, thus avoiding the mode locking seen in the original implementation of the \ac{FPGA} clocked \ac{PD}. Once the measurement interval has completed, the \texttt{done} signal is used to reset the edge detectors asynchronously. It is possible that an edge may occur within the time taken to reset the counters, and thus the detector incorrectly gauge the start of a measurement interval, however this can only occur in exceptional circumstances in the acquisition period. Therefore it is possible to ignore this scenario, as the only impact will be an increased time taken to acquire a lock.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{../new_pdet1}
	\caption[Modified Bang-Bang sign detector RTL diagrams]{Modified Bang-Bang sign detector RTL diagrams.}
	\label{fig:signdet_impl}
\end{figure}

A two signal interface controls the \ac{TDC}, a \texttt{count} that enables the \ac{TDC} and a \texttt{done} signal that signals the termination of the interval. A single logic gate is required to generate each signal, as ORing the two edge detector outputs will produce logic $1$ when the interval has begun while ANDing these signals will indicate the end of the interval when both edges have occurred. The \texttt{done} signal is also used to clock the flip-flop acting as a buffer for the sign. The exact implementation of the sign detection circuit can be seen in Figure \ref{fig:signdet_impl}.

As previously mentioned, the time-to-digital conversion is carried out by a tapped delay line. As with the \ac{RO}, inverters will be used to provide the delay between each tap, and thus the resolution of the \ac{TDL} and the length of each individual delay is based on the layout chosen by the router in Vivado. Due to this, the exact resolution of the phase detector is unknown, although the delay through inverters used in Vivado simulations and the average delay computed based on \ac{RO} centre frequencies can be used to estimate that of the phase detectors. Based these figures for the the propagation delay due to an inverter, $\tau_{inv}$ can be estimated to be in the region of $300~\si{\pico\second}$. As each tap consists of an inverter pair, the delay through the tap, $\tau_{tap} \approx 600~\si{\pico\second}$. Characterisation of the \ac{TDL}, and thus measurement of the delays in each \ac{TDL}, is technically possible, however, this is an time consuming process as many \ac{TDL}s would need to be characterised in order to accurately compute the average delay.

%TODO eugene
The delay line itself is shown in Figure \ref{fig:tdl_impl}, with each tap implemented by a pair of flip-flops and three inverters. For an $N$ bit width phase error detection, $2^{N-1}-1$ taps are required. In the case of this \ac{ADPLL}, a 5 bit two's complement error signal is required, thus the range of the signal is $[-16,15]\cap\mathbb{Z}$. If the range is made symmetrical, 15 taps are required to fill a signal of this size. From the diagram it can be seen that unless edges are detected on each signal at exactly the same instant, it is almost impossible to measure a zero phase difference between reference and generated signals. This is an intentional decision, \cite{idkwhattocite}.
\begin{figure}[h]
	\centering
	\includegraphics[width=1.0\textwidth]{../new_pdet2}
	\caption[Inverter based \ac{TDL} RTL diagrams]{Inverter based \ac{TDL} RTL diagrams.}
	\label{fig:tdl_impl}
\end{figure}
The operation of the \ac{TDL} is as follows: A delayed version of the \texttt{count} signal acts as the clock for flip-flops forming each tap, which are rising edge activated. These flip-flops have a constant ``1'' at their $D$ inputs, so as the signal propagates through the delay line, each flip-flop with propagate a logic ``1'' to their corresponding \texttt{error_tap[i]} signal. When the measurement interval has completed, the \texttt{count} signal clocks the 15 bit wide \texttt{error_tap} signal into \texttt{error_tap_buff} which act as the output buffers of the \ac{PD}, holding the phase error constant between measurement intervals. As with the flip-flops performing edge detection, the flip-flops forming the taps are can only be used once before requiring a reset, which is again carried out using the \texttt[done] signal.

\subsubsection{\acs{ADPLL} 3 Design Summary}
Table \ref{table:adpll3}.

\begin{table}[!h]%TODO numbers!
	\begin{center}
		\begin{tabular}{|l|r|r|r|}
			\multicolumn{4}{c}{\ac{DCO} Tuning Range} \T\\
			\hline
			\multicolumn{1}{|c|}{-} & \multicolumn{1}{c|}{Minimum} & \multicolumn{1}{c|}{Step} & \multicolumn{1}{c|}{Maximum} \T\\
			\hline
			Frequency & $4.571~\si{\mega\hertz}$ & $29.180~\si{\kilo\hertz}$ (at $5~\si{\mega\hertz}$) & $5.518~\si{\mega\hertz}$ \T\\
			\hline
			Period & $218.2~\si{\nano\second}$ & $1.176~\si{\nano\second}$ & $181.1~\si{\nano\second}$ \T\\
			\hline
		\end{tabular}
		\begin{tabular}{|l|r|l|r|}
			\multicolumn{4}{c}{Configuration as Implemented} \T\\
			\hline
			\ac{RO} num. of inverters (max) & 373 & \ac{RO} Control Width & 5 bits \T\\
			\hline
			\ac{RO} Bias Point & 16 & Runtime Gains & Enabled \T\\
			\hline
			Phase Detector Error Width & 5 & Error Sum Weight Width & 4 \T\\
			\hline
			\acs{kp} & $2^{-7}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-9}\times[0,15]\cap\mathbb{Z}$ \T\\
			\hline
			Detection Resolution & $3.875~\si{\nano\second}$ & Detection Phase Resolution & $6.975\si{\degree}$ (at $5~\si{\mega\hertz}$)\\
			\hline
		\end{tabular}
	\end{center}
	\caption[\ac{ADPLL} Design 3 Summary]{\ac{ADPLL} Design 3 Summary.}
	\label{table:adpll3}
\end{table}

\section{Measurement Setup}\label{section:measurement_setup}
For the following sections in which the performance of either individual \ac{ADPLL}s or \ac{ADPLL} networks the same measurement setup was used. The \ac{FPGA} used was a Xilinx Artix-7 \acl{Nexys} on a Digilent \acs{Nexys} board. The \acl{Nexys} features 15850 logic slices, each with 4 lookup tables and 8 flip-flops. There are six clock management tiles, 240 dedicated digital signal processing slices and an analog-to-digital converter. The maximum operating frequency is $464~\si{\mega\hertz}$ and 210 input/output ports are available for use. Further information can be found in the datasheet \cite{a7_datasheet}. The \acs{Nexys} is an evaluation/development board for the aforementioned \ac{FPGA}, however directed at the education sector. It contains useful peripherals such as 16 switches, 8 7-segment displays and 4 user addressable \ac{PMOD} headers and can be used with the free of charge ``Webpack'' version of Vivado \cite{n4_datasheet}.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.6\textwidth]{../n4}
	\caption[\acs{Nexys} development board]{\acs{Nexys} development board.}
	\label{fig:n4}
\end{figure}

Measurements were performed using an Agilent MSO7054A, a four channel, $4~\si[per-mode=symbol]{\giga\sample\per\second}$ oscilloscope. Signals were extracted over the regular \ac{PMOD} headers as the board does not provide outputs better suited to the measurement of signals that are available to the user, such as SMA or BNC connectors, and the \ac{PMOD} header for the analog-to-digital converter, which has lower impedance traces, is not available for the designer to re-configure. Data was acquired with the oscilloscope in quad channel mode at a sampling rate of $2\si{\giga\sample\per\second}$. One of the four channels was consumed by the external reference, leaving three open for signal measurements. $200\si{\micro\second}$ long captures were saved to an Agilent specified binary format which could be opened in Matlab for analysis.
%TODO diagram or picture of some sort

\section{\acs{ADPLL} Characterisation}
%TODO why each PLL
